---
title: "QMH Statistics Club"
author:
- name: "Data Management and Data Summarization"
date:
output:
  prettydoc::html_pretty:
    df_print: paged
    theme: architect
    highlight: github
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    paged.print = FALSE
)
options(knitr.table.format = "html")
```
## Libraries

```{r lib}
library(tidyverse)
library(kableExtra)
library(haven)
```



## Data

Let’s use the given dataset (from Bangladesh Demographic and Health Survey, 2017-18). The respondents were women of reproductive age (15–49 years) who had one or more live birth in the five years preceding the survey.

```{r data}
setwd("C:/Users/has9a/R")
bdhs_data <- read_sav("Data_Competition_level-I.sav")
```

Now take a glimpse of our data by `glimpse(bdhs_data)`.
We see some of the variables are `haven_labelled`. We can convert these labelled variables into factors.
Converting to factor is important for modelling and/or plotting. We will do this by `as_factor` function while solving questions. Also we will check labels of any variable by `attributes()` function. It helps us to find out predefined value for missing values and `NA`



## Discussion
### #Q1

* Selecting required variables for this question and assigning them to another variable.
```{r}
q1 <- bdhs_data %>% select(1,10,11,12)
```

* Observing data by `CASEID`. And checking whether the responses on other variables are same or not for every distinct respondent.
```{r collapse=TRUE}
nrow(distinct(q1)) == nrow(distinct(bdhs_data %>% select(1)))
```

* As it's `TRUE`, we can work with this filtered dataframe (7562 variables).
```{r collapse=TRUE}
q1 <- distinct(q1)
sum(is.na(q1))  # No Missing Values
```

* Now creating new variable `media_exposure`.
```{r}
q1 <- q1 %>%
  mutate(media_exposure = if_else(
    (V157 == 2 | V158 == 2 | V159 == 2 |
       V157 == 3 | V158 == 3 | V159 == 3),
    1, 2
  ))

q1$media_exposure <- labelled(q1$media_exposure,
                             c("Exposed"=1, "Not Exposed"=2), 
                             label="Media Exposure")

q1 = q1 %>%
  mutate(media_exposure = as_factor(media_exposure))
```

* Visualizing `media_exposure`.
```{r}
kable(q1 %>%
  group_by(media_exposure) %>%
  summarise(n = n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 1), "%")),
  col.names = c("Media Exposure", "Frequency", "Percentage"))
```

##### Interpretation

### #Q2

* Selecting required variables for this question and assigning them to another variable.
```{r}
q2 <- bdhs_data %>% select(1,15)
```

* Observing data by `CASEID`. And checking whether the responses on `wealth index factor score` are same or not for every distinct respondent.
```{r collapse=TRUE}
nrow(distinct(q2)) == nrow(distinct(bdhs_data %>% select(1)))
```

* As it's `TRUE`, we can work with this filtered dataframe (7562 variables).
```{r collapse=TRUE}
q2 <- distinct(q2)
sum(is.na(q2))  # No Missing Values
```

* Now creating new variable `economic_status`.
```{r}
q2 <- q2%>%
  mutate( economic_status = cut(V191,3,labels=c("1","2","3")))

q2$economic_status <- labelled(q2$economic_status,
                              c("Poor"=1, "Middle"=2, "Rich"=3), 
                              label="Economic Status")

q2 <- q2 %>%
  mutate(economic_status = as_factor(economic_status))
```

* Visualizing `economic_status`.
```{r}
kable(q2 %>%
  group_by(economic_status) %>%
  summarise(n = n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 1), "%")),
  col.names = c("Economic Status", "Frequency", "Percentage"))
```

##### Interpretation


### #Q3

* Selecting required variables for this question and assigning them to another variable.
```{r}
q3 <- bdhs_data %>% select(1,17,22)
```

* Filtering `q3` by removing first birth and its twins. Adding `NA` where preceding birth interval, `B11`, equals to 999.
```{r}
q3 <- q3 %>% filter(BORD != 1 & B11 != 0)
q3[q3==999] <- NA
```

* As the preceding birth interval data is quantitative, suitable measure of central tendency is either `mean` or `median`.
```{r collapse=TRUE}
summary(q3$B11)
```

* Histogram will help us to to choose better measure of central tendency.
```{r}
hist(q3$B11,breaks=50,
     main = "Histogram To Check Symmetry",
     xlab = "Preceding Birth Interval",
     ylab = "Frequency",
     col = "#ADD8E6")
```
* As it does not seem symmetric, suitable measure of central tendency is `median`.
```{r collapse=TRUE}
median(q3$B11,na.rm = T)
```
##### Interpretation


### #Q4

```{r}
q4 = bdhs_data %>% select(1,5,18,19,20,21)
```

```{r}
q4 <- q4 %>%
  mutate(age_of_child = if_else(
    (B5==1),
    V008-B3, B7
  ))
```

```{r}
q4 <- q4 %>%
  mutate(survival_status = if_else(
    (B5==1),
    "Survived", "Died"
  ))
```

```{r}
hist(q4$age_of_child[q4$B5==0],breaks = 100)
```

Median







